Comparing summed probability distributions of shoreline and radiocarbon dates from the Mesolithic Skagerrak coast of Norway

### 2023-07-28 analysis/paper/paper.Rmd
Discussion

In contrast to the findings of @berg-hansen2022, the analysis undertaken here instead indicates a decline in the number of shoreline dated sites through this time-interval, with possible positive reversals of this trend just after 8000 BCE and around 7000 BCE. It has previously been suggested that a change in mobility patterns occurs from some time around 8300 BCE, involving a reduction in residential mobility [e.g. @damlien2018b; @roalkvam2022]. Furthermore, both genetic data and technological analyses of lithic inventories indicates that the initial phase of human colonisation from the south is followed by an influx, and possible mixing, of people from the east around the same time [@gunther2018; @manninen2021]. If this narrative is coupled with the data observed here, these migratory events could appear to result in a relatively sudden drop in site-count as indicated in the SSPD, which in turn could be related to a change in mobility patterns. The end of this drop also roughly corresponds to the first ^14^C-dates in the RSPD around 8300 BCE -- giving further credence to the suggestion that the appearance of material to be ^14^C-dated is mainly related to cultural and demographic processes as opposed to taphonomic factors.

@@

>>>>

In contrast to the findings of @berg-hansen2022, the analysis undertaken here instead indicates a decline in the number of shoreline dated sites through this time-interval, with possible positive reversals of this trend just after 8000 BCE and around 7000 BCE. It has previously been suggested that a change in mobility patterns occurs from some time around 8300 BCE, involving a reduction in residential mobility [e.g. @damlien2018b; @roalkvam2022]. Furthermore, both genetic data and technological analyses of lithic inventories indicates that the initial phase of human colonisation from the south is followed by an influx, and possible mixing, of people from the east around the same time [@gunther2018; @manninen2021]. If this narrative is coupled with the data observed here, these migratory events could appear to result in a relatively sudden drop in site-count as indicated in the SSPD, which in turn could be related to a change in mobility patterns. The end of this drop also roughly corresponds to the first ^14^C-dates in the RSPD around 8300 BCE -- giving further credence to the suggestion that the appearance of material to be ^14^C-dated is mainly related to cultural and demographic processes.

@@

Given the presence of people and corresponding signal in the SSPD before the start of the RSPD, and the suggested relevance of mobility patterns for the lacking signal in the RSPD in this early period, neither proxy is devoid of influence from population density or mobility patterns. Properly adjudicating this relationship will thus depend on independent data on mobility patterns.
